\section*{Synchronization}
Cooperating Processes execute concurrently, may be interrupted at any time and have access to shared data. Proper sequencing and coordination of processes is needed to maintain data consistency.
% access: LAS or shared memory/message passing
\subsection*{Race Conditions}
\begin{itemize}
    \item \textbf{Execution} outcome dependent on order of concurrent access to shared data
        \item \textbf{examples:} counter in prod-consumer-problem, PID
        \item \textbf{solutions:} disabling interrupts (single core vs. multiprocessors (time-consuming), affects system clock)
        \item \textbf{preemptive kernel}: process preempted in kernel mode, most common. more responsive, suitable for real-time programming
            % TODO: verstehen
    \item \textbf{non-preemptive kernel}: uncommon, process blocks CPU
\end{itemize}
\subsection*{Critical Section Problem}
\begin{itemize}
    \item \textbf{critical section} is where a process accesses shared data
    \item no two processes should be executing in their critical sections at the same time
    \item \textbf{solution:} mutual exclusion, progress (selection cannot be postponed indefinitely), bounded waiting (limit on waiting)
\end{itemize}
\begin{description}
    \item[Peterson's Solution] 2 processes, int turn, bool flag[2] shared, acquire \& lock, not guaranteed to work on modern computer architectures (requires atomic load/store, no instruction reordering)
    \item[Mutex Lock] acquire() and release() lock (atomic), bool available (binary), require busy waiting (spinlock, no context switch required)
    \item[Semaphores] integer variable, accesses through wait() and signal. busy wait \\
        - \textbf{counting} (init to nr of resources available, decr. in wait) vs. \textbf{binary} semaphore (like mutex lock). \\
        -  \textbf{implementation with suspension and waiting queues} %TODO:: S. 26ff
    \item[Monitors] high level form of process sync. \\
        - \textbf{condition variables} wait and signal on condition var.   \\
        - \textbf{mutex locks}
    \item[Priority Inversion] solution: \textbf{priority-inheritance protocol}: inherit higher priority until finished with resources %TODO: Definition  Priority Inversion
    \item[Deadlock]: 2 or more processes are waiting indefinitely for an event that can be caused only by one of the waiting processes.
\end{description}
\subsection*{Examples}
WEITER BEI S. 4 (04 - 2 Examples
\begin{description}
    \item[Bounded Buffer]
    \item[Dining Philosophers]
    \item[Synchronization within the Kernel]
    \item[POSIX Sync]
    \item[Alternative Approaches]
\end{description}
\subsection*{Deadlocks}
% starvation
